---
title: "2020_04_03"
author: "Megan Bontrager"
date: "4/3/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data and libraries:

```{r, warning=FALSE, message=FALSE}
library(rethinking)
library(tidyverse)
library(tidybayes)
library(magrittr)
library(dagitty)

data(bangladesh) 

colnames(bangladesh) = str_replace(colnames(bangladesh), "\\.", "_")

```

Look at data:
```{r}
summary(bangladesh) 
hist(bangladesh$age_centered)
```

Make data into list:

```{r}
dat = list(use_contraception = bangladesh$use_contraception, 
           district = as.numeric(as.factor(bangladesh$district)),
           urban = bangladesh$urban,
           age_centered = bangladesh$age_centered,
           living_children = bangladesh$living_children)
plot(bangladesh$district, dat$district)
```

#### Problem 1

Basic model from problem 1:

```{r, cache = TRUE}
m1 = ulam(
  alist(
    use_contraception ~ dbinom(1, p),
    logit(p) <- a[district] + Bu[district]*urban,
    a[district] ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5)),
  data = dat, chains = 4, log_lik = TRUE, iter = 4e3, cores = 2
  )
```

```{r}
precis(m1, depth = 2)
```

(Plot correlation between intercepts and slopes, plot )

#### Problem 2

Does age affect contraception use directly? or only because of a correlation between age and children?

First, is there a correlation between age and children?

```{r}
ggplot(bangladesh, aes(x = age_centered, y = living_children, color = use_contraception, fill = use_contraception, group = use_contraception)) +
  geom_jitter(aes(color = use_contraception, fill = use_contraception), alpha = 0.2) +
  geom_smooth(method = "lm")
```


Draw dags.

```{r}
dag1 = dagitty("dag { urban -> use
               age -> use
               district -> use
               children -> use }")
drawdag(dag1)

dag2 = dagitty("dag { urban -> use
               age -> children
               district -> use
               children -> use }")
drawdag(dag2)

dag3 = dagitty("dag { urban -> use
               age -> use
               age -> children
               district -> use
               children -> use }")
drawdag(dag3)

```


Age and kids both have separate effects:

```{r, cache = TRUE}

m2 = ulam(
  alist(
    use_contraception ~ dbinom(1, p),
    logit(p) <- a[district] + Bu[district]*urban + Bage*age_centered + Bkids*living_children,
    a[district] ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5),
    Bage ~ dnorm(0, 1.5),
    Bkids ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5)),
  data = dat, chains = 4, log_lik = TRUE, iter = 4e3, cores = 2
  )

```

```{r}
precis(m2)
```


Only age:

```{r, cache = TRUE}

m3 = ulam(
  alist(
    use_contraception ~ dbinom(1, p),
    logit(p) <- a[district] + Bu[district]*urban + Bage*age_centered,
    a[district] ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5),
    Bage ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5)),
  data = dat, chains = 4, log_lik = TRUE, iter = 4e3, cores = 2
  )

```

```{r}
precis(m3)
```

Only kids:

```{r, cache = TRUE}

m4 = ulam(
  alist(
    use_contraception ~ dbinom(1, p),
    logit(p) <- a[district] + Bu[district]*urban + Bkids*living_children,
    a[district] ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5),
    Bkids ~ dnorm(0, 1.5),
    Bu[district] ~ dnorm(0, 1.5)),
  data = dat, chains = 4, log_lik = TRUE, iter = 4e3, cores = 2
  )

```

```{r}
precis(m4)
```

#### Problem 3

Now, try including children as an ordered category, rather than a continuous predictor. 

Level order should be ok. 



```{r, cache = TRUE, eval = FALSE}

# Need to make alpha? 

dat$alpha = rep(1, 3)

m5 = ulam(
  alist(
    use_contraception ~ binomial(1, p),
    logit(p) <- a +
      # a[district] + Bu[district]*urban 
       Bkids*sum(delta_k[1:living_children]),
    a ~ dnorm(0, 1.5),
    # a[district] ~ dnorm(0, 1.5),
    # Bu[district] ~ dnorm(0, 1.5),
    Bkids ~ dnorm(0, 1.5),
    vector[4]: delta_k <<- append_row(0, delta),
    simplex[3]: delta ~ dirichlet(alpha)),
  data = dat, chains = 4, log_lik = TRUE, iter = 4e3, cores = 2
  )


precis(m5, depth = 2, omit = "cutpoints")
```

