---
title: "Chapter16.1"
author: "Megan Bontrager"
date: "8/3/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rethinking)
```

## 16M1

Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model. How do they differ from those of m16.1?

Load and scale data:

```{r}
data("Howell1")
d = Howell1
d$w = d$weight/mean(d$weight)
d$h = d$height/mean(d$height)
```

Original model:

```{r, cache=TRUE, message=FALSE, warning=FALSE, error=FALSE}

m16.1 = ulam(
  alist(
    w ~ dlnorm(mu, sigma), 
    exp(mu) <- 3.141593 * k * p^2 * h^3,
    p ~ beta(2, 18),
    k ~ exponential(0.5),
    sigma ~ exponential(1)
  ),
  data = d, chains = 4, cores = 4
)

```

Look at summary and posterior predictions:

```{r}
precis(m16.1)
pairs(m16.1)
h_seq = seq(from = 0 , to = max(d$w), length.out = 30)
w_sim = sim(m16.1, data = list(h = h_seq))
mu_mean = apply(w_sim, 2, mean)
w_CI = apply(w_sim, 2, PI)
plot(d$h, d$w, xlim = c(0, max(d$h)), ylim = c(0, max(d$w)), col = rangi2, lwd = 2, xlab = "height (scaled)", ylab = "weight (scaled)")
lines(h_seq, mu_mean)
shade(w_CI, h_seq)
```

New model with the exponent on height estimated from the data:

This exponent should be positive, and since we aleardy think it's near 3, going with an exponential on 1, which should have a larger mean than the exp(0.5).

(would be better to do some prior prediction to figure out best prior for l)

```{r, cache=TRUE, message=FALSE, warning=FALSE, error=FALSE}

m16.1b = ulam(
  alist(
    w ~ dlnorm(mu, sigma), 
    exp(mu) <- 3.141593 * k * p^2 * h^l,
    p ~ beta(2, 18),
    k ~ exponential(0.5),
    l ~ exponential(0.1),
    sigma ~ exponential(1)
  ),
  data = d, chains = 4, cores = 4
)

```

Getting some treedepth warnings.

Look at summary and posterior predictions:

```{r}
precis(m16.1b)
pairs(m16.1b)
w_sim = sim(m16.1b, data = list(h = h_seq))
mu_mean = apply(w_sim, 2, mean)
w_CI = apply(w_sim, 2, PI)
plot(d$h, d$w, xlim = c(0, max(d$h)), ylim = c(0, max(d$w)), col = rangi2, lwd = 2, xlab = "height (scaled)", ylab = "weight (scaled)")
lines(h_seq, mu_mean)
shade(w_CI, h_seq)
```

Posterior predictions look similar, but follow the data a bit more at the low end in particular.

## 16M2

Conduct a prior predictive simulation for the cylinder height model. Begin with the priors in the chapter. Do these produce reasonable prior height distributions? If not, which modifications do you suggest?

(Does he mean height or weight?)

```{r}
hist(d$h) # 0 to 1.3
hist(d$w) # 0 to 1.6

pr = extract.prior(m16.1)
mu = link(m16.1, post = pr, data = list(h = h_seq))
summary(exp(mu[,2]))
plot(d$h, d$w, xlim = c(0, max(d$h)), ylim = c(0, max(d$w)), col = rangi2, lwd = 2, xlab = "height (scaled)", ylab = "weight (scaled)")
mu_mean = apply(mu, 2, mean)
w_CI = apply(mu, 2, PI)
lines(h_seq, exp(mu_mean))
shade(exp(w_CI), h_seq)

```

