---
title: "Chapter14.5.2"
author: "Megan Bontrager"
date: "6/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

14M5. Modify the phylogenetic distance example to use group size as the outcome and brain size as a predictor. Assuming brain size influences group size, what is your estimate of the effect? How does phylogeny influence the estimate?

First, look at the phylogeny.
```{r, message=FALSE, error=FALSE}
library(rethinking)
library(knitr)
data(Primates301)
data(Primates301_nex)

library(ape)
plot(ladderize(Primates301_nex), type = "fan", font = 1, no.margin = TRUE, cex = 0.5,label.offset = 1)

```

Prepare data:

```{r}
d = Primates301
d$name = as.character(d$name)
dstan = d[complete.cases(d$group_size, d$body, d$brain),]
spp_obs = dstan$name

dat_list = list(N_spp = nrow(dstan),
                M = standardize(log(dstan$body)),
                B = standardize(log(dstan$brain)),
                G = standardize(log(dstan$group_size)),
                Imat = diag(nrow(dstan)))

```

Prepare cov matrix:

```{r}
tree_trimmed = keep.tip(Primates301_nex, spp_obs)
Rbm = corBrownian(phy = tree_trimmed)
V = vcv(Rbm)
Dmat = cophenetic(tree_trimmed)
plot(Dmat, V)
dat_list$V = V[spp_obs, spp_obs]
dat_list$R = dat_list$V/max(V) 
image(V)
```


Run model with group size as the outcome and brain size as a predictor:

Without phylogeny:

```{r, message = FALSE, cache = TRUE}
m1 = ulam(alist(G ~ multi_normal(mu, SIGMA),
                mu <- a + bB *B +bM*M,
                matrix[N_spp, N_spp]: SIGMA <- Imat * sigma_sq,
                a ~ normal(0, 1),
                c(bM, bB) ~ normal(0, 0.5),
                sigma_sq ~ exponential(1)),
          data = dat_list, chains = 4, cores = 4)
```


```{r}
precis(m1)
plot(precis(m1))
```

With phylogeny:

```{r, message = FALSE, cache = TRUE}
m2 = ulam(alist(G ~ multi_normal(mu, SIGMA),
                mu <- a + bB*B +bM*M,
                matrix[N_spp, N_spp]: SIGMA <- R * sigma_sq,
                a ~ normal(0, 1),
                c(bM, bB) ~ normal(0, 0.5),
                sigma_sq ~ exponential(1)),
          data = dat_list, chains = 4, cores = 4)
```

```{r}
precis(m2)
plot(precis(m2))
```

