---
title: "Chapter9"
author: "Megan Bontrager"
date: "9/27/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(rethinking)
```

## 8E1

Which of the following is a requirement of the simple Metropolis algorithm?
(1)  The parameters must be discrete. <- No
(2)  The likelihood function must be Gaussian. <- No
(3)  The proposal distribution must be symmetric. <- Yes

## 8E2

Gibbs sampling is more efficient than the Metropolis algorithm. How does it achieve this extra efficiency? Are there any limitations to the Gibbs sampling strategy?

Gibbs sampling uses "smarter" proposals to characterize the posterior distribution with fewer samples than a classic metropolis algorithm. It does this with conjugate prior-likelihood pairs, which are analytically tractable.


## 8E3 

Which sort of parameters can Hamiltonian Monte Carlo not handle? Can you explain why?

HMC can't handle discrete parameters, because it can't calculate a gradient across categories. 


## 8M1

Re-estimate the terrain ruggedness model from the chapter, but now using a uniform prior and an exponential prior for the standard deviation, sigma. The uniform prior should be dunif(0,10) and the exponential should be dexp(1). Do the different priors have any detectible influence on the posterior distribution?

```{r, cache=TRUE}

data(rugged)
d <- rugged
d$log_gdp <- log(d$rgdppc_2000)
dd <- d[ complete.cases(d$rgdppc_2000) , ]
dd$log_gdp_std <- dd$log_gdp / mean(dd$log_gdp)
dd$rugged_std <- dd$rugged / max(dd$rugged)
dd$cid <- ifelse( dd$cont_africa==1 , 1 , 2 )

dat_slim <- list(
    log_gdp_std = dd$log_gdp_std,
    rugged_std = dd$rugged_std,
    cid = as.integer( dd$cid )
)
str(dat_slim)

m9.1a <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1 )
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )

m9.1b <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dcauchy( 0 , 1 )
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )

m9.1c <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dunif( 0, 10 )
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )

precis(m9.1a, depth = 2)
precis(m9.1b, depth = 2)
precis(m9.1c, depth = 2)


```

Altering these priors does not have any impact on the estimates of the posterior.


## 8M2

The Cauchy and exponential priors from the terrain ruggedness model are very weak. They can be made more informative by reducing their scale. Compare the dcauchy and dexp priors for progressively smaller values of the scaling parameter. As these priors become stronger, how does each influence the posterior distribution?

```{r, cache=TRUE}
m9.1d <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 1)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )


m9.1e <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 10)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )


m9.1f <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dexp( 100)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )

precis(m9.1d, depth = 2)
precis(m9.1e, depth = 2)
precis(m9.1f, depth = 2)

m9.1g <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dcauchy( 0, 1)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )
m9.1h <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dcauchy( 0, 0.1)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )
m9.1i <- ulam(
    alist(
        log_gdp_std ~ dnorm( mu , sigma ) ,
        mu <- a[cid] + b[cid]*( rugged_std - 0.215 ) ,
        a[cid] ~ dnorm( 1 , 0.1 ) ,
        b[cid] ~ dnorm( 0 , 0.3 ) ,
        sigma ~ dcauchy( 0, 0.01)
),
data=dat_slim , chains=2 , cores=4 , iter=1000 )

precis(m9.1g, depth = 2)
precis(m9.1h, depth = 2)
precis(m9.1i, depth = 2)

```

Varying the exponenetial prior on sigma didn't have a big effect. Same with the Cauchy prior. Maybe I'm not varying them enough?

## 8H1

Run the model below and then inspect the posterior distribution and explain what it is accomplishing.

```{R, cache = TRUE} 
# R code 9.26
mp <- ulam(alist(a ~ dnorm(0,1),
                     b ~ dcauchy(0,1)),
               data=list(y=1),
               start=list(a=0,b=0),
               chains=2 , cores=4 , iter=1000 )

pairs(mp)
traceplot(mp)
```
Compare the samples for the parameters a and b. Can you explain the different trace plots, using what you know about the Cauchy distribution?

The model is sampling from a normal distribution and a cauchy distribution. The Cauchy distribution holds most of it's probability near the mean, but has long fat tails, so it produces a spiky caterpillar. 

## 8H2

Recall the divorce rate example from Chapter 5. Repeat that analysis, using map2stan thistime, fitting models m5.1, m5.2, and m5.3.  Use compare to compare the models on the basis of WAIC. Explain the results.

```{R, cache = TRUE}

data(WaffleDivorce)
d <- WaffleDivorce

# standardize variables
d$A <- scale( d$MedianAgeMarriage )
d$D <- scale( d$Divorce )
d$M <- scale( d$Marriage )

wh = list(A = d$A, D = d$D, M = d$M)


m5.1 <- ulam(alist(D ~ dnorm( mu , sigma ) ,
                   mu <- a + bA * A ,
                   a ~ dnorm( 0 , 0.2 ) ,
                   bA ~ dnorm( 0 , 0.5 ) ,
                   sigma ~ dexp( 1 )) , 
             data = wh, chains=2 , cores=4 , iter=1000, log_lik = TRUE)


m5.2 <- ulam(alist(D ~ dnorm( mu , sigma ) , 
                   mu <- a + bM * M , 
                   a ~ dnorm( 0 , 0.2 ) ,
                   bM ~ dnorm( 0 , 0.5 ) ,
                   sigma ~ dexp( 1 )) ,
             data = wh, chains=2 , cores=4 , iter=1000, log_lik = TRUE)


m5.3 <- ulam(alist(D ~ dnorm( mu , sigma ) ,
                   mu <- a + bM*M + bA*A ,
                   a ~ dnorm( 0 , 0.2 ) ,
                   bM ~ dnorm( 0 , 0.5 ) ,
                   bA ~ dnorm( 0 , 0.5 ) , 
                   sigma ~ dexp( 1 )) ,
             data = wh, chains=2 , cores=4 , iter=1000, log_lik = TRUE)

compare(m5.1, m5.2, m5.3)
```

The model with median age of marriage only is the best.



## 8E4

Explain the difference between the effective number of samples, n_eff as calculated by Stan, and the actual number of samples.

The effective number of samples takes into account the amount of autocorrelation betwee the samples, which is often low when using HMC. As a result, the number of effective samples may be lower than the actual number (if there is some autocorrelation) or could be more than the number of samples (I don't really understand how it ends up being > the actual number of samples).


## 8M3

Re-estimate one of the Stan models from the chapter, but at different numbers of warmup iterations. Be sure to use the same number of sampling iterations in each case. Compare the n_eff values. How much warmup is enough?

```{r, cache = TRUE}

str(dat_slim)

# First with warmup = 500

m9.1.500 <- ulam(
    alist(
        log_gdp_std ~ dnorm(mu, sigma),
        mu <- a[cid] + b[cid]*(rugged_std - 0.215),
        a[cid] ~ dnorm(1 , 0.1) ,
        b[cid] ~ dnorm(0 , 0.3) ,
        sigma ~ dexp(1)
),
data=dat_slim , chains=2 , cores=4 , iter=1000, warmup = 500)

# Now trying 200

m9.1.200 <- ulam(
    alist(
        log_gdp_std ~ dnorm(mu, sigma),
        mu <- a[cid] + b[cid]*(rugged_std - 0.215),
        a[cid] ~ dnorm(1 , 0.1) ,
        b[cid] ~ dnorm(0 , 0.3) ,
        sigma ~ dexp(1)
),
data=dat_slim , chains=2 , cores=4 , iter=700, warmup = 200)

# Now 50

m9.1.50 <- ulam(
    alist(
        log_gdp_std ~ dnorm(mu, sigma),
        mu <- a[cid] + b[cid]*(rugged_std - 0.215),
        a[cid] ~ dnorm(1 , 0.1) ,
        b[cid] ~ dnorm(0 , 0.3) ,
        sigma ~ dexp(1)
),
data=dat_slim , chains=2 , cores=4 , iter=550, warmup = 50)

# Now 10
m9.1.10 <- ulam(
    alist(
        log_gdp_std ~ dnorm(mu, sigma),
        mu <- a[cid] + b[cid]*(rugged_std - 0.215),
        a[cid] ~ dnorm(1 , 0.1) ,
        b[cid] ~ dnorm(0 , 0.3) ,
        sigma ~ dexp(1)
),
data=dat_slim , chains=2 , cores=4 , iter=510, warmup = 10)


precis(m9.1.500, depth = 2)
precis(m9.1.200, depth = 2)
precis(m9.1.50, depth = 2)
precis(m9.1.10, depth = 2)

```

Even a warmup of just 50 iterations produces effective sample sizes > 200. But dropping down to 10 results in some warnings: 

WARNING: No variance estimation is performed for num_warmup < 20

There were 2 chains where the estimated Bayesian Fraction of Missing Information was low. See http://mc-stan.org/misc/warnings.html#bfmi-lowExamine the pairs() plot to diagnose sampling problems

Need to figure out some terminology: iteration, transition, sample. Are these all the same, i.e., 500 iterations after warmp = 500 transitions and results in 500 samples?


## 8H3

Sometimes changing a prior for one parameter has unanticipated effects on other parameters. This is because when a parameter is highly correlated with another parameter in the posterior, the prior influences both parameters. Here’s an example to work and think through.

Go back to the leg length example in Chapter 5. Here is the code again, which simulates height and leg lengths for 100 imagined individuals:

```{r}
# R code 9.26
N <- 100
height <- rnorm(N,10,2)
leg_prop <- runif(N,0.4,0.5)
leg_left <- leg_prop*height +
    rnorm( N , 0 , 0.02 )
leg_right <- leg_prop*height +
    rnorm( N , 0 , 0.02 )
d <- data.frame(height,leg_left,leg_right)
# number of individuals
# sim total height of each
# leg as proportion of height
# sim left leg as proportion + error
# sim right leg as proportion + error
# combine into data frame
plot(d$leg_left, d$leg_right)
# Very correlated, good for standina and walking.
```

And below is the model you fit before, resulting in a highly correlated posterior for the two beta parameters. This time, fit the model using map2stan:

```{r, cache = TRUE}
# R code 9.27
m5.8s <- ulam(
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + bl*leg_left + br*leg_right,
        a ~ dnorm(10, 100),
        bl ~ dnorm(2, 10),
        br ~ dnorm(2, 10),
        sigma ~ dcauchy(0, 1)
),
data = d, chains = 4, start = list(a = 10, bl = 0, br = 0, sigma = 1), log_lik = TRUE)

```

Compare the posterior distribution produced by the code above to the posterior distribution produced when you change the prior for br so that it is strictly positive:

```{r, cache = TRUE}
# R code 9.29
m5.8s2 <- map2stan(
    alist(
        height ~ dnorm(mu, sigma),
        mu <- a + bl*leg_left + br*leg_right,
        a ~ dnorm(10, 100),
        bl ~ dnorm(2, 10),
        br ~ dnorm(2, 10) & T[0, ],
        sigma ~ dcauchy(0, 1)
),
data = d, chains = 4, start = list(a = 10, bl = 0, br = 0, sigma = 1), log_lik = TRUE)
```

Note that T[0,] on the right-hand side of the prior for br. What the T[0,] does is truncate the normal distribution so that it has positive probability only above zero. In other words, that prior ensures that the posterior distribution for br will have no probability mass below zero.

Compare the two posterior distributions for m5.8s and m5.8s2. What has changed in the posterior distribution of both beta parameters? Can you explain the change induced by the change in prior?

```{r}
precis(m5.8s, depth = 2)

precis(m5.8s2, depth = 2)

pairs(m5.8s)
pairs(m5.8s2)

```

By constraining the right leg parameter to be positive, it means that the left leg parameter is sometimes more negative than when the right leg is unconstrained, so the posterior for the left leg has a longer negative tail and the right leg parameter has a longer positive tail. 


## 8H4

For the two models fit in the previous problem, use DIC or WAIC to compare the effective numbers of parameters for each model. Which model has more effective parameters? Why?

```{r, cache = TRUE}

WAIC(m5.8s) # pWAIC = 4.06
WAIC(m5.8s2) # pWAIC = 3.69

```

The number of effective parameters is slightly lower for the model with the constrained prior on the right leg. Perhaps this is because that parameter is less "unknown" in the second model, so it doesn't contribute as much to the variance in log-probability?

